### 2013-09-04 methods.tex

Methods & Materials
Study Inclusion Criteria & MR Imaging
Prostates removed by radical prostatectomy were used in this IRB-approved (Duke IRB# Pro00006458), HIPAA compliant study from men ranging in age from XX--XX diagnosed with biopsy-proven prostate cancer destined for surgical removal. Between DATES, a total of XX patients were recruited and enrolled in this study. Inclusion criteria were undergoing complete pelvic MRI with endorectal coil for detection of prostate cancer, including multiplanar T2-weighted anatomic imaging, diffusion-weighted imaging (DWI), and dynamic contrast enhanced MRI (DCE-MRI) as well as radical prostatectomy and whole mount histology. Patients with previous treatments of prostate cancer or benign prostatic hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded.  All patients enrolled in this study provided written informed consent. Our final cohort included XX subjects. Table XX has a summary of the patient demographics for the cases shown in this manuscript.
All imaging was performed on one of two 3.0 Tesla MR scanners (General Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare, Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad, Indianola, PA) as well as multichannel surface coils.  Imaging sequences included thin-section (3 mm section thickness) fast spin echo T2-weighted images in the coronal, axial and sagittal planes.  Diffusion weighted images were obtained using multiple b-values and calculation of ADC maps was also performed.  Dynamic contrast enhanced MR sequences were obtained after administration of a weight-based dose of extracellular MR contrast agent with 4-5 second temporal resolution for 5-6 minutes. (If we need it, can put in table with full MR parameters)
Prostates were radically removed using a da Vinci Surgical System.  After exision, the prostates were formalin fixed for at least 24 hours without being cut, and then processed for whole mount histology.
Whole Mount Histology
Fixed in formalin; hand-cut ~3 mm spacing
H&E whole-mount images
Clinically read, with the following markings (ELN)
Digitally-scanned @ 600 dpi
Manually segmented using itk-SNAP; saved as nifti stack
Rigid Histology Image Registration
MR T2WI Data
Diffeomorphic MR:Histology Image Registration
Rigid T2WI:ADC Image Registration