### 2013-09-13 methods.tex
Stephen ARFI methods

\input{mr_methods}

\input{arfi_methods}

\input{image_proc_methods}