### 2013-09-13 acknowledgements.tex
updated formatting

\section*{ACKNOWLEDGEMENTS}
The authors would like to thank Siemens Medical Solution USA, Ultrasound
Division for their in-kind technical support and Ned Danieley for system administration support.
This work was supported by NIH R01CA142824 and the Duke Coulter Translational Grant Program.