### 2013-09-13 background.tex

\section{BACKGROUND}\label{sect:background}

\input{mr_background}