### 2013-09-13 methods.tex

\section{METHODS}\label{sect:methods}

\input{mr_methods}

\input{arfi_methods}

\input{image_proc_methods}