### 2013-09-13 background.tex
updated new refs and anatomy background

\section{Background}\label{sect:background}

\input{prostate_anatomy}

\input{mr_background}