### 2013-09-13 tab_mr_arfi_volumes.tex
updated formatting in table and figs

\begin{table}
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Zone} & {\bf ARFI Total} \\ 
& {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline
\input{data/vol_data}
\hline
\end{tabular}
\label{tab:mr_arfi_volumes}
\end{table}