### 2013-09-15 tab_mr_arfi_axes.tex
redid presentation of mr/arfi axis table data

\begin{table}
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Major / Minos Axes in MR T2WI and ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|l|l|} \hline
{\bf Study} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} & {\bf MR} & {\bf ARFI} \\ 
{\bf Subject} & {\bf CA} & {\bf CA} & {\bf CL} & {\bf CL} & {\bf CE} & {\bf CE} & {\bf TA} & {\bf TA} & {\bf TL} & {\bf TL} & {\bf TE} & {\bf TE} \\
 & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} & {\bf (cm)} \\ \hline
\input{data/axis_data}
\hline
\end{tabular}
\label{tab:mr_arfi_axes}
\end{table}