### 2013-09-15 fig_mr_arfi_central_axes.tex
added central axis mr:arfi fig

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes1} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes2} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_axes3} \\
(a) Lat-Lat : Apex-Base & (b) Ant-Post : Apex-Base & (c) Lat-Lat : Ant-Post \\
\end{tabular}
\label{fig:mr_arfi_central_axes} 
\end{figure}