### 2013-09-16 tab_axis_ratio_over_under.tex

\begin{table}
\centering
\caption{Mean axis ratio differences between ARFI imaging, T2WI MR and
    pathology for the total prostate volume and central gland / zone for the
    imaging modalities.  The three axis ratios analyzed were:
    lateral-to-lateral : apex-to-base (LL:AB), anterior-to-posterio :
    apex-to-base (AP:AB), and lateral-to-lateral : anterior-to-posterior
    (LL:AP).}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Image Modality} & {\bf Comparative Measure} & {\bf Total / Central} & {\bf Axes} & {\bf Axis Ratio Difference (\%)} \\ \hline
\input{figs/tab_axis_ratio_over_under_data}
\hline
\end{tabular}
\label{tab:axis_ratio_over_under}
\end{table}