### 2013-09-21 fig_mr_segs_vol.tex
P79 seg vol added

\begin{figure}[htb!]
\centering
\includegraphics[width=1.0\textwidth]{figs/MR_P79_SegsVol.eps}
\caption{Representative MR T2WI segmentations and rendered volume performed in
    3D Slicer, with the peripheral zone being delineated in red, and the
    central gland being delineated in blue.  The native imaging plane for
    segmentation is the axial view, shown in the bottom left image.  The bottom
    middle and right images show the projections of the rendered model segment
    outlines in the sagital and coronal views, respectively.}
\label{fig:mr_segs_vol} 
\end{figure}