### 2013-09-22 fig_mr_arfi_path_axes.tex
added error estimates for mr:arfi axes

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/Lateral-to-Lateral} &
\includegraphics[width=0.3\linewidth]{figs/Anterior-to-Posterior} &
\includegraphics[width=0.3\linewidth]{figs/Apex-to-Base} \\
(a) & (b) & (c) \\
\includegraphics[width=0.3\linewidth]{figs/Imaging_Lateral-to-Lateral} &
\includegraphics[width=0.3\linewidth]{figs/Imaging_Anterior-to-Posterior} &
\includegraphics[width=0.3\linewidth]{figs/Imaging_Apex-to-Base} \\
(d) & (e) & (f) \\
\end{tabular}
\caption{Measurements of the prostate dimensions along the three standard
    anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
    apex-to-base (c).  The correlation between the MR and ARFI imaging axis
    measurements was performed for each orientation.  The black dashed-line
    represents the projection of perfectly-correlated measurements between
    imaging and patholoy.  Notice that there is no correlation between imaging
    and pathology in the lateral-to-lateral and anterior-to-posterior axes,
    with a general over-estimation of the lateral-to-lateral axis in the
    imaging.  The MR estimation of the apex-to-base dimension showed moderate
    correlation with pathology, with the ARFI image apex-to-base approximation
    having an overall underestimation in this orientation.  The axis
    measurements of the imaging modalities realtive to one another, for both
    total gland (black) and central gland (red) are shown in the bottom row.
    The over-/under-estimation of each imaging modality relative to gross
    pathology and each other is summarized in
    Table~\ref{tab:mr_arfi_path_axes}.} 
\label{fig:mr_arfi_path_axes}
\end{figure}