### 2013-09-22 fig_mr_arfi_volumes.tex
integrated lin reg vol figs

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from 19.6--71.0
    cm$^3$ based on MR image models (a), with a moderate correlation between MR
    and ARFI imaging for both the total gland volume (R$^2$ = 0.68, (b)) and the
    central gland (R$^2$ = 0.41, (c)).  The ARFI total gland volumes had a positive bias
    compared with the MR volumes (36.7 $\pm$ 27.9\% (b)), while the ARFI
    central zone volumes were on average more similar (2.1 $\pm$ 39.0\%, (c)), but
    with some significant variability between study subjects.
    Table~\ref{tab:mr_arfi_volumes} contains the individual volume estimates
    for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}