### 2013-09-22 background.tex

\section{Background}\label{sect:background}

\input{prostate_anatomy}

\input{clinical_significance}

\input{mr_background}

\input{arfi_background}