### 2013-09-22 path_methods.tex

\subsection{Pathology Processing}
After excision, the prostates were weighed and tri-axially measured, formalin
fixed for at least 24 hours without being cut, and then processed for whole
mount histology.  While tri-axial measurements of the gross prostate were not necessarily anatomically aligned, rough estimates of the prostate volume were made using a tri-axial, ellipsoidal organ volume assumption, where the volume could be estimated as:

\begin{equation}
V = \frac{4}{3}\pi \sqrt{\det{A^{-1}}},
\end{equation}

where $A$ represents the eigenvector matrix of the tri-axial pathology
measurements.