### 2013-09-22 study_methods.tex

\subsection{Study Inclusion Criteria}
Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
HIPAA-compliant study.  A total of 16 patients were recruited and enrolled in
this study.  Inclusion criteria were undergoing complete pelvic MRI with
endorectal coil for detection of prostate cancer, including multiplanar
T2-weighted anatomic imaging, as well as pre-operative ARFI imaging and radical
prostatectomy.  Patients with previous treatments of PCa or benign prostatic
hyperplasia (BPH), or anatomic anomalies of the rectum, were excluded from this
study.  All patients enrolled in this study provided written informed consent.

\emph{CO-AUTHOR QUESTION: DO WE WANT TO INCLUDE A TABLE OF PATIENT DEMOGRAPHIC INFORMATION?}