### 2013-10-07 path_methods.tex
re-organized radical prostatectomy text

\subsection{Pathology Processing}
Prostates were radically removed using a da Vinci Surgical System (Intuitive
Surgical\textregistered, Inc., Sunnyvale, CA).  After excision, the prostates
were weighed and tri-axially measured, formalin fixed for at least 24 hours
without being cut, and then processed for whole mount histology.  While
tri-axial measurements of the gross prostate were not necessarily anatomically
aligned, rough estimates of the prostate volume were made using a tri-axial,
ellipsoidal organ volume assumption, where the volume could be estimated as:

\begin{equation}\label{eqn:ellipsoid_volume}
V = \frac{4}{3}\pi \sqrt{\det{A^{-1}}},
\end{equation}

where $A$ represents the eigenvector matrix of the tri-axial pathology
measurements.