### 2013-10-07 fig_mr_arfi_volumes.tex
updated text to reflect new volume data (Fig 4)

\begin{figure}[htb!]
\centering
\begin{tabular}{ccc}
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_volumes} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_total_linreg} &
\includegraphics[width=0.3\linewidth]{figs/mr_arfi_central_linreg} \\
(a) CG:PZ Volume & (b) Total Gland Volume & (c) Central Gland Volume \\
\end{tabular}
\caption{Comparison of MR and ARFI zonal anatomy volume estimates from
    manually-segmented images.  Total prostate gland volumes ranged from
    19.6--71.0 cm$^3$ based on MR image models (a), with a moderate correlation
    between MR and ARFI imaging for both the total gland volume (R$^2$ = 0.63,
    (b)) and the central gland (R$^2$ = 0.38, (c)).  The ARFI total gland
    volumes had a positive bias compared with the MR volumes (6.1 $\pm$ 25.0\%
    (b)), while the ARFI central zone volumes were on average more similar
    (-5.0 $\pm$ 39.5\%, (c)), but with some significant variability between
    study subjects.  Table~\ref{tab:mr_arfi_volumes} contains the individual
    volume estimates for the total prostate and central glands.}
\label{fig:mr_arfi_volumes} 
\end{figure}