### 2013-10-07 coverletter.tex
coverletter

\documentclass[10pt]{article}
\usepackage{graphicx}
\pagestyle{empty}
\topmargin=-0.75in
\oddsidemargin=0.0in
\textwidth=6.5in
\textheight=9.5in
\parindent=0.0in
\parskip=10pt

\begin{document}
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\includegraphics[width=3.0in,trim=55 50 50 85,clip=true]{/home/mlp6/bme/Duke_BME_Logos/Pratt-HorizontalLogo-Vector-Black}
\end{minipage}\hfill
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\hfill{\bf Biomedical Engineering}\\

\begin{small}
\hfill 136 Hudson Hall, Box 90281, Durham, NC 27708-0281

\hfill Tel: (919) 660 5158 $\bullet$ Fax: (919) 684 4488

\hfill mark.palmeri@duke.edu\\
\end{small}
\end{minipage}
\begin{Large}

\vspace{0.25in}
\today
\vspace{0.25in}

Re: Ultrasonic Imaging Manuscript Submission
\vspace{0.25in}

To whom it may concern:
\vspace{0.1in}

We thank you for taking the time and effort to review our manuscript, titled
{\it ``Acoustic Radiation Force Impulse (ARFI) Imaging Prostate Zonal Anatomy:
    Comparison with Endorectal T2-Weighted MR Imaging (T2WI)''}, for
consideration to be published in Ultrasonic Imaging.  All of the authors on
this manuscript have reviewed its content and approve of its submission.  This
work has not been published or submitted for review to any other journals.

Please do not hesitate to contact me if you have any questions regarding this
submission.

\vspace{0.25in}

%\hspace*{0.575\textwidth} Sincerely,
Sincerely,\\
\includegraphics{/home/mlp6/MarkLPalmeriSig}\\
%\vspace*{0.5in}
Mark L. Palmeri, M.D., Ph.D.\\
\end{Large}
Assistant Research Professor\\
Biomedical Engineering \& Anesthesiology
%\hspace*{0.575\textwidth} Mark L. Palmeri, M.D., Ph.D.\\
%\hspace*{0.575\textwidth} Assistant Research Professor\\
%\hspace*{0.575\textwidth} Biomedical Engineering \& Anesthesiology
\end{document}