### 2013-12-09 conclusions.tex
updated volume Rsq and errors to Matlab vars

\section{Conclusions}
The delineation of prostate zonal anatomy in ARFI images has been compared with
the established methods for identifying zonal anatomy using MR T2W images.
Both imaging modalities showed moderate correlations between estimated organ
volume and gross pathologic weights, and ARFI:MR total prostate gland volumes
were well-correlated (R$^2$ = \MRarfiVolTotalRsq), but ARFI images yielded
prostate volumes that were, on average, larger (\MRarfiVolTotalMeanDiff\%~$\pm$
\MRarfiVolTotalStdDiff\%) than MR images, primarily due to over-estimation of
the lateral dimension of the prostate total gland (\ARFImrTotalLatLatMeanPct~
$\pm$ \ARFImrTotalLatLatStdPct\%), while over-estimates of the other dimensions
were less significant contributors (\ARFImrTotalAntPostMeanPct~$\pm$
\ARFImrTotalAntPostStdPct\%~and \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\%).  The central zone volumes of ARFI and MR images
were also well-correlated (R$^2$ = \MRarfiVolCentralRsq), with minimal volume
bias between the imaging modalities, but significant variability case-to-case
(\MRarfiVolCentralMeanDiff~$\pm$ MRarfiVolCentralStdDiff\%).  Central zone
volume differences were, again, strongly attributed to over-estimation of the
lateral dimension (\ARFImrCentralLatLatMeanPct~$\pm$
\ARFImrCentralLatLatStdPct\%), with a significant underestimation of the
anterior-to-posterior dimension (\ARFImrCentralAntPostMeanPct~$\pm$
\ARFImrCentralAntPostStdPct\%).  Strong variability in central gland volumes is
believed to be related to the extent of benign prostatic hyperplasia (BPH) for
select cases.  Overall, ARFI imaging of the prostate yielded prostate volumes
and dimensions that were correlated with MR T2WI estimates, with biases in the
anterior-to-posterior dimension, most likely related to poor displacement SNR
in the anterior region of the prostate from greater distance from the rectal
wall imaging surface, and the lateral dimension, where contrast between the PZ
and peri-prostatic fat could be limited.  ARFI imaging is a promising low-cost,
real-time imaging modality that can compliment MR imaging for diagnosis,
treatment planning and management of PCa.