### 2013-12-29 study_methods.tex
minor text updates

\subsection{Study Inclusion Criteria}
Patients undergoing radical prostatectomy for biopsy-proven PCa treatment were
enrolled as study subjects in this IRB-approved (Duke IRB\# Pro00006458),
HIPAA-compliant study, with all enrolled patients providing written informed
consent.  A total of 16 patients were recruited and enrolled in this study.
Inclusion criteria were undergoing complete pelvic MRI with endorectal coil for
detection of prostate cancer, including multiplanar T2-weighted anatomic
imaging, as well as pre-operative ARFI imaging and radical prostatectomy.
Patients with previous treatments of PCa or benign prostatic hyperplasia (BPH),
or anatomic anomalies of the rectum, were excluded from this study.