### 2013-12-29 mr_methods.tex
minor text edits

\subsection{MR Imaging}
All MR imaging was performed on one of two 3.0 Tesla MR scanners (General
Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare,
Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
Indianola, PA), as well as multi-channel surface coils.  Imaging sequences
included thin-section (3 mm section thickness) fast spin echo T2-weighted
images in the coronal, axial and sagittal planes.  Diffusion weighted images
were obtained using multiple b-values and calculation of Apparent Diffusion
Coefficient (ADC) maps was also performed.  Dynamic contrast enhanced MR
sequences were obtained after administration of a weight-based dose of
extracellular MR contrast agent with 4-5 second temporal resolution for 5-6
minutes.