### 2014-01-03 discussion.tex

\section{Discussion}\label{sect:discussion}
\input{disc_arfi_mr}

\input{limitations}

\input{future}