### 2014-01-03 fig_setup_annotated.tex
new annotated setup figure

\begin{figure}[htb!]
\centering
\includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
\caption{The experimental setup used for ARFI imaging in the operating suite,
    including the Siemens SC2000 scanner with the ER7B endorectal transducer
    sitting in a custom cradle of a CIVCO Micro-Touch\texttrademark stabalizer
    with optical feedback controlling a stepper motor to acquire 3D ultrasound
    imaging data \invivo.  Custom-written Python code was run on the scanner to
    trigger B-mode and ARFI imaging acquisition sequences relative to completed
    rotation increments of $\sim$ 1-2\degree.}
\label{fig:setup_annotated} 
\end{figure}