### 2014-01-06 disc_arfi_mr.tex
disc_arfi_mr.tex edited online with Bitbucket

\subsection{ARFI and MR T2WI Agreement}
This study demonstrated that ARFI imaging delineates zonal anatomy of the
prostate well when compared to MR T2WI with only slight overestimation of 
ARFI total and central gland volume (Figure~\ref{fig:mr_arfi_volumes}). 
 Establishing accurate delineation of the prostate's zonal anatomy is 
important because PCa has strong correlations with location in the prostate, 
specifically in the peripheral zone, and accurate capsule delineation will f
acilitate future efforts to register 3D imaging datasets between ARFI and MR imaging.

We did not expect perfect agreement between MR and ARFI imaging by volume or tri-axial
 dimensions since each modality applies a different amount of external compression to 
the rectal wall that deforms the prostate different amounts. Pressure on the prostate 
during ARFI imaging is modulated by the urologist when he locks the transducer rotation 
apparatus in place prior to imaging, and deformation of the prostate can be easily visualized
 during this alignment process.  It was our goal to apply as little pressure while still maintaining
 uniform contact throughout the image acquisition. This is critical since poor acoustic coupling to 
the prostate through the rectal wall can greatly diminish ARFI image displacement SNR.

The lateral boundaries of the prostate in the U/S dataset were overestimated relative to MR T2WI, increasing 
ARFI total volume relative to MR. The lateral overestimation was a result of poor contrast between the PZ and 
peri-prostatic fat in the B-mode images, and image decimation along the lateral dimension in both ARFI and B-mode 
images. Without image decimation, the elevation and axial orthogonal planes can be used to correct errors introduced
 by the poor contrast between the PZ and peri-prostatic fat. However, with image decimation, segmentation changes in 
the elevation and axial orthogonal planes were limited to the lower bound on the model precision of, again, $\sim$3 mm.
 While reduced decimation could improve the precision of the lateral measurements, this would have introduced significant 
overhead in the manual segmentation process, with diminishing returns given the spatial smoothing of the modeling algorithms
 that were utilized.

The anterior aspect of the prostate can also be challenging to delineate,
especially in large prostates, where SNR decreases with increasing depth away
from our rectal wall imaging surface and MR images suffer from gland
heterogeneity in this region~\cite{Gupta2013}, though we had good agreement
between ARFI and MR T2W images in the anterior-to-posterior dimension
(Figure~\ref{fig:mr_arfi_path_axes}(b), Table~\ref{tab:mr_arfi_axes_error}).

ARFI imaging consistently underestimated the extent of the prostate capsule and 
central gland apex-to-base dimension relative to MR T2WI ((Figure~\ref{fig:mr_arfi_path_axes}(c), 
Table~\ref{tab:mr_arfi_axes_error}).  In fact, it is likely the degree of ARFI estimation relative 
to the actual apex-to-base prostate length is even greater than suggested by the MR comparison as this 
was the dimension with the worst spatial resolution in MR imaging, with a
typical MR T2WI slice thickness of $\sim$3 mm for these studies.  While
higher-resolution MR images could be obtained and analyzed for more precise
comparisons in this imaging dimension, such data was not available for this
study.