### 2014-01-20 results.tex

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with moderate correlation
between the ARFI and MR total prostate gland volumes (R$^2$ =
\MRarfiVolTotalRsq), with a mean overestimation of of
\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ARFI imaging compared
to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland volumes
were well-correlated between ARFI and MR images (R$^2$ = \MRarfiVolCentralRsq)
with no significant mean over/under-estimation (\MRarfiVolCentralMeanDiff~$\pm$
\MRarfiVolCentralStdDiff\%), but significant variability between the cases
(Figure~\ref{fig:mr_arfi_volumes}(c)).  Table~\ref{tab:mr_arfi_volumes} has the
individually-measured volumes for MR and ARFI imaging for each study subject.

\input{fig_mr_arfi_volumes}

Weights and axis measurements from the gross pathology processing of the
excised prostates were collected (Table~\ref{tab:path_data}), and using the
axis measurements (lateral-to-lateral, anterior-to-posterior, and
apex-to-base), the prostate volume was approximated as a tri-axial ellipsoid,
and its volume was estimated (\ref{eqn:ellipsoid_volume}).  Prostate weights
were moderately correlated with estimated pathology ellipsoidal prostate
volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = \pathVolWeightRsq).  There
was moderate correlation between the prostate weight and the
image-reconstructed prostate volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$
= \weightMRrsq~(MR) and \weightARFIrsq~ (ARFI)), though there was weaker
correlation with the ellipsoidal approximation of the measurement prostate
volume and the image-reconstructed volumes (Figure~\ref{fig:mr_arfi_weight}(c),
R$^2$ = \pathVolMRrsq~(MR) and \pathVolARFIrsq~(ARFI)).

\input{fig_mr_arfi_weight}

Measurements of the prostate total gland and CG dimensions along the three
standard anatomic axes (apex-to-base, lateral-to-lateral, and
anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
correlation between the imaging axis measurements was analyzed
(Figure~\ref{fig:mr_arfi_path_axes}).  ARFI was most correlated to MR in the
lateral-to-lateral axis in both the total gland and CG (R$^2$ =
\totalLatLatRsq~and \centralLatLatRsq, respectively), with mean overestimates
of \ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  ARFI had moderate correlation with the
total prostate gland and CG axes in the anterior-to-posterior dimension (R$^2$
= \totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  ARFI imaging
had weak correlation with MR images in the apex-to-base dimension (R$^2$ =
\totalApexBaseRsq, total gland and R$^2$ = \centralApexBaseRsq, CG), with
underestimates of \ARFImrTotalApexBaseMeanPct~$\pm$
\ARFImrTotalApexBaseStdPct\% and \ARFImrCentralApexBaseMeanPct~$\pm$
\ARFImrCentralApexBaseStdPct\%, respectively.

\input{fig_mr_arfi_path_axes}

\input{tab_mr_arfi_axes_error}

%\input{fig_mr_arfi_total_axes}

%\input{fig_mr_arfi_central_axes}

%\input{tab_axis_ratio_over_under}