### 2014-01-21 abstract.tex

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second leading cause of cancer-related death.  Two
imaging modalities, Multi-parametric Magnetic Resonance Imaging (mpMRI) and
Acoustic Radiation Force Imaging have the potential to aid in PCa diagnosis and
management by evaluating the structural composition of prostate zones and
tumors based on water content and stiffness respectively.  In this study,
\invivo MR and ARFI imaging datasets were compared to one another and against
gross pathology in order to evaluate their ability to delineate central gland
(CG) and prostate capsule anatomy. 3D slicer, an open source image processing
program,  was used for MR and ARFI manual for manual segmentation, 3d modeling,
and measurement of zonal anatomy tri-axial dimensions and volumes.   Both
imaging modalities showed moderate correlations between modeled organ volume
and gross pathologic weights.  In the intra-modality comparison, ARFI and MR
total prostate gland volumes were well-correlated (R$^2$ = \MRarfiVolTotalRsq),
but ARFI yielded prostate volumes that were, on average, larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR. This was
primarily due to total gland over-estimation along the lateral dimension.
(\ARFImrTotalLatLatMeanPct $\pm$ \ARFImrTotalLatLatStdPct\%). Differences
between MR and ARFI in the other dimensions were less significant
(\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
\ARFImrTotalApexBaseStdPct\%, apex-to-base). ARFI and MR CG volumes were also
well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
attributed to MR apex-to-base axis underestimation
(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
lateral dimension over-estimation  (\ARFImrCentralLatLatMeanPct $\pm$
\ARFImrCentralLatLatStdPct\%).  Overall, ARFI prostate imaging yielded prostate
volumes and dimensions that were correlated with MR T2WI estimates, with biases
in the lateral dimension. ARFI imaging is a promising low-cost, real-time
imaging modality that can compliment mpMRI for diagnosis, treatment planning
and management of PCa.