### 2014-01-21 future.tex

\subsection{Future Directions}
Future directions for our research include deeper analysis of our findings,
especially the lateral dimension over-estimation, to improve our anatomic
delineation and assess further its clinical impact.  Technological improvements
to improve anterior prostate boundary delineation in both B-mode and ARFI
imaging are also being pursued for better visualization in this region.  3D
models of the prostate and central glands will be used across ARFI and MR
imaging datasets to spatially register them to facilitate correlation of
regions of PCa suspicion between the two modalities, and to correlate with
whole-mount histology of the excised prostate specimens.  This effort will
allow the sensitivity and specificity of each imaging modality for PCa
detection and characterization to be quantified.  Additionally, we will be
pursuing quantitative measurement of the pressure being applied during imaging
to help reduce prostate deformation and reduce the impact of prostate stiffness
nonlinearity in our displacement images.  While ARFI imaging of the prostate is
a constantly evolving technology, the fact that one can view the different
prostatic zones on ARFI images in addition to MRI is encouraging and could lead
to targeted diagnostic biopsies and therapies with real-time imaging for men
with PCa.