### 2014-02-08 abstract.tex
shortened abstract for TMI limits

\section*{Abstract}
Prostate cancer (PCa) is the most common non-cutaneous malignancy among men in
the United States and the second-leading cause of cancer-related death.
Multi-parametric Magnetic Resonance Imaging (mpMRI) has gained recent
popularity to characterize PCa.  ARFI imaging has the potential to aid PCa
diagnosis and management by using tissue stiffness to evaluate prostate zonal
anatomy and lesions.  MR and ARFI \invivo imaging datasets were compared to one
another and with gross pathology measurements made immediately after radical
prostatectomy.  Images were manually segmented in 3D Slicer to delineate the
central gland (CG) and prostate capsule, and 3D models were rendered to
evaluate zonal anatomy dimensions and volumes.  Both imaging modalities showed
good correlation between estimated organ volume and gross pathologic weights.
Ultrasound and MR total prostate volumes were well-correlated (R$^2$ =
\MRarfiVolTotalRsq), but ARFI images yielded prostate volumes that were larger
(\MRarfiVolTotalMeanDiff\% $\pm$ \MRarfiVolTotalStdDiff\%) than MR images, due
to overestimation of the lateral dimension (\ARFImrTotalLatLatMeanPct $\pm$
\ARFImrTotalLatLatStdPct\%), with less significant differences in the other
dimensions (\ARFImrTotalAntPostMeanPct $\pm$ \ARFImrTotalAntPostStdPct\%,
anterior-to-posterior, and \ARFImrTotalApexBaseMeanPct $\pm$
\ARFImrTotalApexBaseStdPct\%, apex-to-base).  ARFI and MR CG volumes were also
well-correlated (R$^2$ = \MRarfiVolCentralRsq).  CG volume differences were
attributed to ARFI underestimation of the apex-to-base axis
(\ARFImrCentralApexBaseMeanPct $\pm$ \ARFImrCentralApexBaseStdPct\%), and ARFI
overestimation of the lateral dimension (\ARFImrCentralLatLatMeanPct $\pm$
\ARFImrCentralLatLatStdPct\%).  ARFI imaging yielded prostate volumes and
dimensions that were well-correlated with MR T2WI estimates, with biases in the
lateral dimension due to poor contrast caused by peri-prostatic fat.  ARFI
imaging is a promising low-cost, portable, real-time modality that can
compliment mpMRI for PCa diagnosis, treatment planning and management.