### 2014-02-08 acknowledgements.tex
modified format and cover letter for Ultrasonic Imaging submission

\section*{Acknowledgements} 
The authors would like to thank Siemens Medical Solution USA, Ultrasound
Division for their in-kind technical support.  Special thanks to Ned Danieley
for computer system administration support, the pathology assistants and Dr.
John Madden in the Duke University Medical Center for their gross prostate
organ processing, and Matthew Brown for his assistance in designing the optical
prostate rotation feedback system.  This work was supported by NIH R01CA142824
and the Duke Coulter Translational Grant Program.