### 2014-02-08 coverletter.tex

\documentclass[10pt]{article}
\usepackage{graphicx}
\pagestyle{empty}
\topmargin=-0.75in
\oddsidemargin=0.0in
\textwidth=6.5in
\textheight=9.5in
\parindent=0.0in
\parskip=10pt

\newcommand{\journal}{Ultrasonic Imaging}
\begin{document}
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\includegraphics[width=3.0in,trim=55 50 50 85,clip=true]{/home/mlp6/Documents/Duke_BME_Logos/Pratt-HorizontalLogo-Vector-Black}
\end{minipage}\hfill
\begin{minipage}[t]{0.5\linewidth}
\vspace{0pt}
\hfill{\bf Biomedical Engineering}\\

\begin{small}
\hfill 136 Hudson Hall, Box 90281, Durham, NC 27708-0281

\hfill Tel: (919) 660 5158 $\bullet$ Fax: (919) 684 4488

\hfill mark.palmeri@duke.edu\\
\end{small}
\end{minipage}
\begin{Large}

\vspace{0.25in}
\today
\vspace{0.25in}

Re: \journal~Manuscript Submission
\vspace{0.25in}

To whom it may concern:
\vspace{0.1in}

We thank you for taking the time and effort to review our manuscript, titled
{\it ``Acoustic Radiation Force Impulse (ARFI) Imaging Prostate Zonal Anatomy:
    Comparison with 3T T2-Weighted MR Imaging (T2WI)''}, for consideration to
be published in \journal.

All of the authors on this manuscript have reviewed its content and approve of
its submission.

This work has not been published in or submitted for review to any other
journals.

The human subject data presented in this manuscript was acquired in a Duke
University Medical Center IRB-approved study, with all study subjects providing
written informed consent.

Please do not hesitate to contact me if you have any questions regarding this
submission.

\vspace{0.25in}

%\hspace*{0.575\textwidth} Sincerely,
Sincerely,\\
\includegraphics{/home/mlp6/Documents/MarkLPalmeriSig}\\
%\vspace*{0.5in}
Mark L. Palmeri, M.D., Ph.D.\\
\end{Large}
Assistant Research Professor\\
Biomedical Engineering \& Anesthesiology
%\hspace*{0.575\textwidth} Mark L. Palmeri, M.D., Ph.D.\\
%\hspace*{0.575\textwidth} Assistant Research Professor\\
%\hspace*{0.575\textwidth} Biomedical Engineering \& Anesthesiology
\end{document}