### 2014-04-23 mr_methods.tex

\subsection{MR Imaging}
All MR imaging was performed on one of two 3.0 Tesla MR scanners (General
Electric HDx, GE Healthcare, Waukesha, WI;  Siemens Skyra, Siemens Healthcare,
Erlangan, Germany) using a single channel Medrad eCoil endorectal coil (Medrad,
Indianola, PA), as well as multi-channel surface coils.  Imaging sequences
included thin-section (3 mm section thickness) fast spin echo T2-weighted
images in the coronal, axial and sagittal planes.  Diffusion weighted images
(DWI) were obtained using multiple b-values and calculation of Apparent
Diffusion Coefficient (ADC) maps was also performed.  Dynamic contrast enhanced
MR (DCE-MRI) sequences were obtained after administration of a weight-based
dose of extracellular MR contrast agent \textbf{(Magenvist\textregistered,
Bayer Pharma AG)} with 4-5 second temporal resolution for 5-6 minutes.
\textbf{\input{mr_dwi_dce_methods}}