### 2014-04-24 fig_setup_annotated.tex

\begin{figure}
\centering
\includegraphics[width=0.75\textwidth]{figs/setup_annotated.png}
\caption{The experimental setup used for \textbf{B-mode/}ARFI imaging in the
    operating suite, including the Siemens SC2000 scanner with the ER7B
    endorectal transducer sitting in a custom cradle of a CIVCO
    Micro-Touch\texttrademark stabilizer with optical feedback controlling a
    stepper motor to acquire 3D ultrasound imaging data \invivo.
    Custom-written Python code was run on the scanner to trigger B-mode and
    ARFI imaging acquisition sequences relative to completed rotation
    increments of $\sim$1\degree.}
\label{fig:setup_annotated} 
\end{figure}