### 2014-04-24 results.tex

\section{Results}\label{sect:results}
The subvolumes associated with the zonal anatomy in each imaging modality were
measured (Figure~\ref{fig:mr_arfi_volumes}(a)), with good correlation between
the ultrasound and MR total prostate gland volumes (R$^2$ =
\MRarfiVolTotalRsq), with a mean overestimation of of
\MRarfiVolTotalMeanDiff~$\pm$ \MRarfiVolTotalStdDiff\% by ultrasound imaging
compared to MR volumes (Figure~\ref{fig:mr_arfi_volumes}(b)).  Central gland
volumes were well-correlated between ARFI and MR images (R$^2$ =
\MRarfiVolCentralRsq) with no significant mean over/under-estimation
(\MRarfiVolCentralMeanDiff~$\pm$ \MRarfiVolCentralStdDiff\%), but significant
variability between the cases (Figure~\ref{fig:mr_arfi_volumes}(c)).
Table~\ref{tab:mr_arfi_volumes} has the individually-measured volumes for MR
and \textbf{B-mode/}ARFI imaging for each study subject.

\input{fig_mr_arfi_volumes}

Weights and axis measurements from the gross pathology processing of the
excised prostates were collected (Table~\ref{tab:path_data}), and using the
axis measurements (lateral-to-lateral, anterior-to-posterior, and
apex-to-base), the prostate volume was approximated as a tri-axial ellipsoid,
and its volume was estimated (\ref{eqn:ellipsoid_volume}).  Prostate weights
were moderately correlated with estimated pathology ellipsoidal prostate
volumes (Figure~\ref{fig:mr_arfi_weight}(a), R$^2$ = \pathVolWeightRsq).  There
was moderate correlation between the prostate weight and the
image-reconstructed prostate volumes (Figure~\ref{fig:mr_arfi_weight}(b), R$^2$
= \weightMRrsq~(MR) and \weightARFIrsq~ (\textbf{B-mode})), though there was weaker
correlation with the ellipsoidal approximation of the measurement prostate
volume and the image-reconstructed volumes (Figure~\ref{fig:mr_arfi_weight}(c),
R$^2$ = \pathVolMRrsq~(MR) and \pathVolARFIrsq~(\textbf{B-mode})).

\input{fig_mr_arfi_weight}

Measurements of the prostate total gland and CG dimensions along the three
standard anatomic axes (apex-to-base, lateral-to-lateral, and
anterior-to-posterior) were made (Table~\ref{tab:mr_arfi_axes}), and the
correlation between the imaging axis measurements was analyzed
(Figure~\ref{fig:mr_arfi_path_axes}). \textbf{Ultrasound images were} most
correlated to MR in the lateral-to-lateral axis in both the total gland
\textbf{B-mode} and CG \textbf{(ARFI)} (R$^2$ = \totalLatLatRsq~and
\centralLatLatRsq, respectively), with mean overestimates of
\ARFImrTotalLatLatMeanPct~$\pm$ \ARFImrTotalLatLatStdPct\%~and
\ARFImrCentralLatLatMeanPct~$\pm$ \ARFImrCentralLatLatStdPct\%, respectively
(Table~\ref{tab:mr_arfi_axes_error}).  \textbf{Ultrasound images} had moderate
correlation with the total prostate gland \textbf{(B-mode)} and CG
\textbf{ARFI}) axes in the anterior-to-posterior dimension (R$^2$ =
\totalAntPostRsq~and \centralAntPostRsq), respectively, with differences of
\ARFImrTotalAntPostMeanPct~$\pm$ \ARFImrTotalAntPostStdPct\% and
\ARFImrCentralAntPostMeanPct~$\pm$ \ARFImrCentralAntPostStdPct\%.  \textbf{The
ultrasound images} had weak correlation with MR images in the apex-to-base
dimension (R$^2$ = \totalApexBaseRsq, total gland and R$^2$ =
\centralApexBaseRsq, CG), with underestimates of
\ARFImrTotalApexBaseMeanPct~$\pm$ \ARFImrTotalApexBaseStdPct\% and
\ARFImrCentralApexBaseMeanPct~$\pm$ \ARFImrCentralApexBaseStdPct\%,
respectively.

\input{fig_mr_arfi_path_axes}

\input{tab_mr_arfi_axes_error}

%\input{fig_mr_arfi_total_axes}

%\input{fig_mr_arfi_central_axes}

%\input{tab_axis_ratio_over_under}