### 2014-04-24 fig_mr_arfi_path_axes.tex
updated caption to reflect usage of B-mode for total gland

\begin{figure}
\centering
\begin{tabular}{ccc}
%\includegraphics[width=0.3\linewidth]{figs/Lateral-to-Lateral} &
%\includegraphics[width=0.3\linewidth]{figs/Anterior-to-Posterior} &
%\includegraphics[width=0.3\linewidth]{figs/Apex-to-Base} \\
%(a) & (b) & (c) \\
\includegraphics[width=0.3\linewidth]{figs/Imaging_Lateral-to-Lateral} &
\includegraphics[width=0.3\linewidth]{figs/Imaging_Anterior-to-Posterior} &
\includegraphics[width=0.3\linewidth]{figs/Imaging_Apex-to-Base} \\
(a) & (b) & (c) \\
%(d) & (e) & (f) \\
\end{tabular}
\caption{Measurements of the prostate dimensions along the three standard
    anatomic axes: lateral-to-lateral (a), anterior-to-posterior (b) and
    apex-to-base (c).  The correlation between the MR and B-mode/ARFI imaging
    axis measurements was performed in each orientation for the total gland
    (red crosses) and central gland (blue circles).  \textbf{B-mode images were
        used for total gland measurements, and ARFI images were for central
        gland measurements.} The black dashed-line represents the projection of
    perfectly-correlated measurements between imaging and pathology.  The
    over-/under-estimation of each imaging modality relative to gross pathology
    and each other is summarized in Table~\ref{tab:mr_arfi_axes_error}.} 
\label{fig:mr_arfi_path_axes}
\end{figure}