### 2014-04-24 tab_mr_arfi_volumes.tex

\begin{table}[h!]
\centering
\caption{Comparison of Central Gland / Zone and Total Prostate Volumes in MR T2WI and \textbf{B-mode/}ARFI Imaging}
\begin{tabular}{|l|l|l|l|l|} \hline
{\bf Study Subject} & {\bf MR Central Gland} & {\bf MR Total} & {\bf ARFI Central Gland} & {\bf B-mode Total} \\ 
& {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} & {\bf Volume (cm$^3$)} \\ \hline
\input{data/vol_data}
\hline
\end{tabular}
\label{tab:mr_arfi_volumes}
\end{table}