### 2014-04-24 arfi_resolution.tex
final responses in place; still need to spell check and proof read

\textbf{The first-order limiting parameters in our ultrasound spatial
resolution to segment capsules and central glands were: (1) the scan conversion
of our native radial ultrasound imaging planes to yield uniform 3D voxels of
B-mode and ARFI imaging data, and (2) the decimation of imaging volume datasets
for manual segmentation.  The scan-conversion interpolation artifact is worse
at deeper depths, where native image data is more sparsely available.  For
these reasons, neither the native imaging plane resolution of B-mode or ARFI
imaging is considered a limiting factor in the total and central gland
measurements presented in this work.}