### 2014-04-24 disc_arfi_mr.tex

\subsection{B-mode/ARFI and MR T2WI Agreement}
This study demonstrated that \textbf{B-mode/}ARFI imaging delineates zonal
anatomy of the prostate well when compared to MR T2WI with only slight
overestimation of \textbf{B-mode} total and \textbf{ARFI} central gland volume
(Figure~\ref{fig:mr_arfi_volumes}).  Establishing accurate delineation of the
prostate's zonal anatomy is important because PCa has strong correlations with
location in the prostate, specifically in the peripheral zone, and accurate
capsule delineation will facilitate future efforts to register 3D imaging
datasets between ARFI and MR imaging.

\input{disc_deformation}

The lateral boundaries of the prostate in the \textbf{ultrasound} dataset were
overestimated relative to MR T2WI, increasing \textbf{B-mode} total volume
relative to MR. The lateral overestimation was a result of poor contrast
between the PZ and peri-prostatic fat in the B-mode images, and image
decimation along the lateral dimension in both ARFI and B-mode images. Without
image decimation, the elevation and axial orthogonal planes can be used to
correct errors introduced by the poor contrast between the PZ and
peri-prostatic fat. However, with image decimation, segmentation changes in the
elevation and axial orthogonal planes were limited to the lower bound on the
model precision of, again, $\sim$3 mm.  While reduced decimation could improve
the precision of the lateral measurements, this would have introduced
significant overhead in the manual segmentation process, with diminishing
returns given the spatial smoothing of the modeling algorithms that were
utilized.

\input{arfi_resolution}

The anterior aspect of the prostate can also be challenging to delineate,
especially in large prostates, where SNR decreases with increasing depth away
from our rectal wall imaging surface and MR images suffer from gland
heterogeneity in this region~\cite{Gupta2013}, though we had good agreement
between ultrasound and MR T2W images in the anterior-to-posterior dimension
(Figure~\ref{fig:mr_arfi_path_axes}(b), Table~\ref{tab:mr_arfi_axes_error}).

\textbf{B-mode/}ARFI imaging consistently underestimated the extent of the
prostate capsule and central gland apex-to-base dimension relative to MR T2WI
((Figure~\ref{fig:mr_arfi_path_axes}(c), Table~\ref{tab:mr_arfi_axes_error}).
In fact, it is likely the degree of \textbf{B-mode/}ARFI underestimation
relative to the actual apex-to-base prostate length is even greater than
suggested by the MR comparison as this was the dimension with the worst spatial
resolution in MR imaging, with a typical MR T2WI slice thickness of $\sim$3 mm
for these studies.  While higher-resolution MR images could be obtained and
analyzed for more precise comparisons in this imaging dimension, such data were
not available for this study.