### 2014-04-24 path_volume_choice.tex
added Buck reference on formalin prostate contraction

An alternative approach to estimating the prostate volume was to reconstruct
the prostate capsule from serial whole mount histology slides of the prostate.
This approach, however, was not chosen because the formalin fixation process
can cause 0-25\% contraction of the tissue~\cite{Schned1996}, the $\sim$3 mm
slice thickness introduces error, and tearing / folding of tissue on the whole
mount slides introduce additional artifacts that need to be compensated for in
the volume estimation process.  Finally, differences in the tissue orientation
on each whole mount slide would require an image registration process to align
all of the slides, introducing another source of error in the process.