### 2014-04-25 image_proc_methods.tex
updated response on extraprostatic tissue based on Buck feedback

\subsection{Image Zonal Anatomy Segmentation and 3D Model Rendering}
\subsubsection{MR Image Segmentation and Modeling}
Axial MR T2WI images were manually segmented by a \textbf{single radiologist
    (CK)} using the smooth polygon tool in ITK-SNAP
(http://www.itksnap.org)~\cite{Yushkevich2006}, using unique labels for the PZ,
CG and AFS. The gland was segmented from base-to-apex.  The base was identified
below the bladder, and subsequent images were segmented until the last slice
with visible prostatic tissue was identified caudally. The CG, urethra, PZ and
AFS were segmented independently according to their well-established anatomical
characteristics on
T2WI.~\cite{Barentsz2012,Jung2012,Poon1985,Hricak2007,Bonekamp2011} The PZ was
identified by its homogeneous high signal intensity on T2WI, which is usually
similar to that of the nearby extraprostatic fat. The CG was visualized and
delineated based on its heterogeneous, lower signal intensity and its location
(Figure~\ref{fig:mr_anatomy}). The urethra was included in the CG segmentation.
When visible, the AFS was identified by its low T2 signal intensity and its
location anterior to the CG.  \textbf{All MR image segmentations were
    ultimately reviewed by an experienced prostate radiologist (RTG).}

\input{fig_mr_anatomy}

Segmented image stacks were imported into 3D Slicer (http://www.slicer.org, v4.3.0)~\cite{Fedorov2012} and 3D
models were rendered using the following parameters (Table~\ref{tab:3dslicer}):

\begin{table}[h!]
\centering
\caption{3D model volume rendering parameters}
\begin{tabular}{ll}
{\bf Parameter} & {\bf Value} \\ \hline
Decimation & 0.1 \\
Smoothing Algorithm & Laplacian \\
Smoothing  & 70.0 \\
Joint Smoothing & Enabled \\
\end{tabular}
\label{tab:3dslicer}
\end{table}

The 3D models were used to render volume estimates of the PZ and CG, with the
sum of PZ and CG representing the total prostate gland volume. AFS volume
estimates in select MR cases were included in the total prostate gland volume
estimates. Orthogonal tri-axial measurements in the lateral-to-lateral,
apex-to-base, and anterior-to-posterior dimensions were made in 3D Slicer by
placing an ROI cube in the model's center and expanding the ROI cube to fit the
maximal model dimensions (Figure~\ref{fig:roi_tool}).

\input{fig_roi_tool}

Figure~\ref{fig:mr_anatomy} shows the anatomic zones and some anatomic
orientation labels in the prostate of a representative study subject MR T2WI.
The zonal anatomy of the prostate was manually segmented
(Figure~\ref{fig:mr_segs_vol}, bottom left) across the image stack, and 3D
models of the zonal anatomy volumes were rendered
(Figure~\ref{fig:mr_segs_vol}, top).

\input{fig_mr_segs_vol}

\subsubsection{Ultrasound Image Segmentation and Modeling}
Sagittal B-mode and ARFI ultrasound image stacks were segmented in 3D Slicer
\textbf{by two researchers (TJG \& ZAM)} using the \verb+draweffect+ tool to
model and measure the prostate capsule and CG volume and tri-axial dimensions
(Figure~\ref{fig:arfi_segs}).  \input{bmode_arfi_coregistered} The sagittal imaging plane was chosen for
segmentation because it has a greater number of easily identifiable anatomical
structures:  seminal vesicles in the left and right posterior base, urethra in
the mid apex, surrounding fibromuscular tissue in the anterior, and rectum in
the posterior. Prostate capsule was segmented in B-mode as this modality offers
both high contrast between prostatic tissue and surrounding extraprostatic fat,
and better SNR in the anterior capsule region than ARFI images. ARFI images
were used for CG segmentation.  Prior to segmentation, both the B-mode and ARFI
image stacks were down-sampled along the sagittal axis to a slice thickness of
3 mm to reduce the number of manually-segmented frames (18 vs. 360 slices /
image stack) and match the MRI slice thickness.

While segmentations were performed along the sagittal plane, label map
intersections in the axial and coronal views were used during active creation
of the sagittal segmentation to further improve segmentation accuracy.
Parameters used for 3D model rendering were identical to those used for MR
modeling (Table~\ref{tab:3dslicer}).  The modeled capsule helped guide the CG
segmentation by enabling capsule model intersections in the three orthogonal
planes during CG segmentation (Figure~\ref{fig:arfi_segs}). The capsule
provided an outer boundary for the CG segmentation. The capsule and CG models
dimensions were then measured using the same procedure as the MR images.
\textbf{All B-mode and ARFI image segmentations were reviewed by an experienced
    ARFI imaging researcher (KRN).}

\input{fig_arfi_segs}