### 2014-05-02 authors.tex
broke out authors

\author{
    \IEEEauthorblockN{
        Mark L. Palmeri, M.D., Ph.D.\IEEEauthorrefmark{1}, 
        Zachary A. Miller\IEEEauthorrefmark{1}, 
        Tyler J. Glass\IEEEauthorrefmark{1}, 
        Kirema Garcia-Reyes\IEEEauthorrefmark{2}, 
        Rajan T. Gupta, M.D.\IEEEauthorrefmark{3}, 
        Stephen J.  Rosenzweig\IEEEauthorrefmark{1}, 
        Christopher Kauffman, M.D.\IEEEauthorrefmark{3}, 
        Thomas J. Polascik, M.D.\IEEEauthorrefmark{4},
        Andrew Buck, M.D.\IEEEauthorrefmark{5},
        Samantha L. Lipman\IEEEauthorrefmark{1},
        Ned C. Rouze\IEEEauthorrefmark{1},
        Kathryn R. Nightingale, Ph.D.\IEEEauthorrefmark{1}}

    \IEEEauthorblockN{
        {\IEEEauthorrefmark{1}Department of Biomedical Engineering, Pratt School of Engineering, Duke University}
        {\IEEEauthorrefmark{2}Duke University School of Medicine}
        {\IEEEauthorrefmark{3}Department of Radiology, Duke University Medical Center}
        {\IEEEauthorrefmark{4}Department of Surgery (Urology), Duke University Medical Center}
        {\IEEEauthorrefmark{5}Department of Pathology, Duke University Medical Center}
    }
}