### 2014-05-02 arfi_methods.tex
corrected small typos

\subsection{B-mode/ARFI Imaging}
Experimental \textbf{B-mode and} ARFI imaging data were acquired using a modified Siemens Acuson
SC2000\texttrademark~ultrasound scanner (Siemens Healthcare, Ultrasound Business Unit,
Mountain View, CA, USA) and the \textbf{side-fire array of an Acuson ER7B
endorectal transducer}.  The ARFI imaging sequence was comprised of
standard B-mode ultrasonic imaging, or tracking beams, and pushing beams. For
each lateral location, two pre-push reference images were acquired, then three
300 cycle pushing pulses were transmitted in rapid succession, focused at 30
mm, 22.5 mm, and 15 mm, respectively, and finally the response of the tissue
was tracked for up to 6 ms at a PRF of 8 kHz. This pushing strategy is similar to
what has been published by Bercoff \etal~\cite{Bercoff2004}. The 30 mm and 22.5
mm foci pushing pulses were transmitted at 4.6 MHz with a F/2 geometry, and the
15 mm focus pushing pulse was transmitted at 5.4 MHz with a F/2.35 geometry to
maintain the same beamwidth (0.67 mm) throughout the region of excitation. A
total of 82 lateral locations were interrogated to cover the 55 mm field of
view, translating 0.67 mm laterally per location.

For the tracking pulses, 16 parallel receive lines at 5.0 MHz were spaced to
observe the response of the tissue to the pushing pulses.  Specifically, four
lines were dedicated to tracking the on-axis displacement, with all 4 beams
located inside the beamwidth of the pushing pulses such that the tracking beam
spacing was 0.17 mm.  The tracking beams were focused at 60 mm with F/3
transmit and F/0.5 receive focal configurations.

\textbf{3D volumetric imaging data were acquired endorectally} using mechanical
rotation of the ER7B transducer between sequential imaging frames, sweeping
across the lateral extent of the prostate.  This rotation setup utilized a
CIVCO Micro-Touch\texttrademark stabilizer (CIVCO Medical Solutions, Kalona, IA
USA) that allows for complete 6-axis degrees of freedom for manual positioning
of the transducer to sweep through the entire prostate during imaging.
Sequential imaging frames were $\sim$1\degree apart, depending on the absolute
size of the prostate.  A custom optical angular feedback transduction circuit
utilizing a reflective linear strip with 212 lines-per-inch (LPI) resolution
(US Digital, Vancouver, WA, USA) with a QSB-S Quadrature-to-USB adapter to
achieve 9 line / degree resolution.  Rotation was performed with a 141 oz-in
torque stepper motor with planetary gearbox (Model \#11YPG202S-LW4-R27, Anaheim
Automation, Anaheim, CA, USA) (Figure~\ref{fig:setup_annotated}) to achieve
accurate spatial localization of the imaging frames in the 3D dataset.

\input{fig_setup_annotated}

Displacement estimation was performed using a phase-shift estimator on the
beamformed in-phase and quadrature (IQ) data~\cite{Loupas95,pinton06}.
\input{arfi_normalization}

The normalized displacement data were scan-converted using a 2D Delaunay
triangulation interpolation method in Matlab (R2012b, The MathWorks, Inc,
Natick, MA, USA) for each lateral location in the dataset.  The scan-converted
ARFI imaging data had a final isotropic voxel size of 0.15 $\times$ 0.15
$\times$ 0.15 mm$^3$.