### 2014-05-02 arfi_normalization.tex

\textbf{The ARFI data were normalized as a function of depth to account for   
attenuation and focal gain effects to ensure that ARFI image brightness
differences were more directly related to stiffness differences.  This                
normalization was performed using a displacement profiles measured in two            
homogeneous, elastic tissue-mimicking phantoms (CIRS, Norfolk, VA) with a          
Youngs modulus of 10.0 kPa and acoustic attenuations of 0.49 and 0.72 dB/cm/MHz.        
The mean displacement profiles from each phantom were averaged together to
achieve a mean attenuation of 0.6 dB/cm/MHz, which is within the measured
range of attenuation in healthy peripheral zone prostate
tissue~\cite{Parker1993}.  This mean displacement profile was low-pass filtered
with a cutoff-frequency of 0.8 mm$^{-1}$ and applied to all displacements in
the entire dataset at each time step.}